Prelim Outline

### 2019-06-13 draft.tex
reduced to 13 pages! references and maybe bigger/fewer figures?

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage[margin=1.0in]{geometry}
\usepackage{float}
\usepackage{ulem}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{wrapfig}
\renewcommand{\baselinestretch}{1.5}

\begin{document}
\section{Introduction}
A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. In the forefront of modern research are materials like high temperature superconductors and heavy fermion systems for which simple effective theories have not yet been found. For example, the cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space  with a robust anti-ferromagnetic (AFM) charge-transfer insulator state near zero doping, a prominent superconducting dome in the 0.1 - 0.2 hole doped region at low temperatures with psuedo-gap and non-Fermi liquid metallic phases at higher temperatures and Fermi liquid like behavior in the overdoped area, of which only the AFM insulating and Fermi liquid phases have well understood mechanism and models. The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures. As in the cuprate case, models for these phases have not been fully fleshed out. Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging.

The fundamental barrier in writing down low-energy theories for strong correlated materials is the existence of non-perturbative interactions between various degrees of freedom - for example, lattice, spin and charge. Consider just the cuprate case for now, the two phases which are well understood contain a singular important degree of freedom in the low-energy space. Near zero doping the cuprates are charge-transfer insulators with an optical gap around 2 eV , indicating that the lowest energy excitations are primarily spin-like, allowing for a simple nearest-neighbor spin-only model like an AFM Heisenberg model to describe the low-energy physics quite accurately. In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles , indicating our lowest energy excitations are electron-like, and a simple non-interacting electron model (with renormalized masses, etc) will suffice. In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where most difficulties arise in writing down low-energy effective models. On top of that, this entire discussion did not include any lattice effects like structural phase transitions which many strongly-correlated materials have as well.

@@

We propose using a density matrix downfolding (DMD) procedure to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials. The Hubbard model is a simple model with nearest-neighbor hopping and a local on-site Coulomb repulsion, parameterized by a single ratio U/t: 
\begin{equation}
H_\text{hub} = -t \sum_{\langle i,j \rangle,\sigma}( c_{i^\dagger,\sigma} c_{j,\sigma} + h.c.) + U \sum_i n_{i\uparrow} n_{i\downarrow}
\label{hubbard}.
\end{equation}
The solutions in 1-d are well understood via the Bethe ansatz , however the 2-d square lattice Hubbard model has no known exact solution.  In the very strong coupling limit $U/t \rightarrow \infty$ an emergent low-energy spin-space exists with an effective nearest-neighbor AFM Heisenberg Hamiltonian. In the region $U = 0$ a trivial non-interacting theory holds. Between these two limits, in particular the intermediate coupling region $2 < U < 8$, no well established effective theory exists. In this region there is a metal-insulator transition at $U_c/t \sim 4$ , and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$. In addition to the half-filling phase diagram, this model exhibits a plethora of interesting ground state properties when doped including stripe phases and d-wave pairing. The relative simplicity of the Hubbard model and the lack of a well-established low-energy theory in a broad intermediate coupling region lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems.

The DMD procedure allows us to develop low-energy effective theories for physical systems in a systematic manner beginning with the \textit{ab-initio} Hamiltonian:
\begin{equation}
\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|}
\label{eq:Hab}
\end{equation}
where $i$ indicates electrons and $I$ nuclei, and the unit of energy is Hartree (Ha).
In this paper we will be working under the Born-Oppenheimer approximation and core electrons will be treated using pseudo-potentials.
The procedure defines a method for downfolding the \textit{ab-initio} energy functional and Hilbert space onto an effective energy functional defined on a low-energy subspace of the Hilbert space 
\begin{equation}
(E_\text{ab}, \mathcal{H}) \xrightarrow{\text{DMD}} (E_\text{eff}, \mathcal{LE} \subset \mathcal{H}), \text{ s.t. }
\forall |\Psi\rangle \in \mathcal{LE}, \ E_\text{eff}[\Psi] - E_\text{ab}[\Psi] = \epsilon[\Psi] + E_0.
\label{eq:DMD}
\end{equation} 
The energy functionals can be written as expectations over Hamiltonian operators $E_\text{ab} = \langle \Psi | \hat{H}_\text{ab} |\Psi \rangle$, $E_\text{eff} = \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$, $\epsilon$ represents the error in our effective theory, and $E_0$ is a constant energy shift.
Broadly the DMD method consists of four steps which will be discussed in detail later: 
A) Defining the space of low-energy excitations which $E_\text{eff}$ should be able to describe accurately.
B) Sampling, at minimum, a set of states in $\mathcal{H}$ which probe the chosen low-energy excitations in a statistically independent manner. The selected states need not be eigenstates. 
C) Constructing a set of candidate models which are linear in reduced density matrix (RDM) descriptors with variable coefficients $\{c\}$
\begin{equation}
E_\text{eff} = \sum_k c_k \langle \Psi | \hat{d}_k |\Psi \rangle + E_0.
\label{eq:Eeff}
\end{equation}
Here $\hat{d}_k$ are Hermitian operators typically composed of 1- or 2-RDM elements. 
Examples of possible $\hat{d}_k$ are number operators $\hat{n}_i$ and exchange operators $\vec{S}_i \cdot \vec{S}_j$.
Additionally, a single particle basis must be built to express RDM elements on and should be chosen such that variations among states in $\mathcal{LE}$ are mostly described by variations in the projection of these states onto the basis.
D) Using the samples to fit the coefficients $\{c\}$ by linear regression such that $E_\text{eff}$ satisfies the conditions in \eqref{eq:DMD} with the smallest error $\epsilon$ possible. 
The fitting procedure is typically an ordinary linear regression, however, one may use a custom cost function in the regression as well. 
In order to conduct the linear regression for a functional like \eqref{eq:Eeff}, one needs to calculate for each state sampled from $\mathcal{LE}$ the expectation values of $\hat{H}_\text{ab}$ and every descriptor $\{\hat{d}_k\}$ used in the regression.
In principle, if conducted with no approximations, the DMD method is rigorously guaranteed to obtain exact results, namely $\epsilon[\Psi] = 0$.

When downfolding $\hat{H}_\text{ab}$ we will use FN-DMC to access states in $\mathcal{LE}$ and accurately calculate their \textit{ab-initio} energies and RDM elements.
Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function.
Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}_\text{ab}$ with ground state $|\Phi_0\rangle$. We apply the projector $e^{-\tau \hat{H}_\text{ab}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
\begin{equation}
\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} |\Psi_T\rangle 
\equiv \lim_{\tau \rightarrow \infty} |\Psi_\text{DMC}(\tau)\rangle \propto \langle \Phi_0|\Psi_T\rangle |\Phi_0\rangle,
\end{equation}
projecting out the ground state as long as the trial wave function we choose is not orthogonal to the ground state. 
The stochastic implementation involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H}_\text{ab} - E_T)} | R^\prime \rangle$. Since $H_\text{ab} = T + V$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(V(R) + V(R^\prime))/2} \langle R| e^{-d\tau(\hat{T} - E_T)}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.
This implementation, however, suffers from a fermion sign problem. 
We deal with this sign problem through a fixed-node approximation, where the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}_\text{ab}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.
We make use of this variational principle to access low-energy states within $\mathcal{H}$ by choosing trial wave functions with nodes that differ from the nodes of $|\Phi_0 \rangle$.
To calculate the expectation value of an operator $\hat{A}$ in FN-DMC we will use a mixed estimator $\langle \Psi_{DMC}(\tau) |\hat{A} | \Psi_T \rangle/\langle \Psi_\text{DMC}(\tau) | \Psi_T \rangle$ which allows us to use the importance sampled distribution $f = \Psi_\text{DMC}(\tau)\Psi_T$ for our Monte Carlo estimates.
Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.

@@

\subsection{Effective theory for CuO molecule using DMD}
As a first step towards developing accurate models for extended systems like solids using DMD, we constructed a many-body effective model for the CuO molecule which accurately describes the eigenstates and spectra seen in experiment up to 2eV above the ground state.
The neutral CuO molecule has been a subject of intense theoretical and experimental study due to the complex structure of its low-energy space.
While the low-energy excitations in the doublet sector of the CuO molecule are well understood, singlet selection rules for spectroscopic measurements and the doublet ground state leave the quartet sector of excitations unexplored.
We construct our $\mathcal{LE}$ based on recent anion photoelectron spectroscopy (APES) measurements of the low-lying excited states of the nuetral CuO molecule. 
The lowest energy excitations measured, which lie below 2eV, involve primarily the O 2p$_z$, 2p$_\pi$ and Cu 4s orbitals, which are fully-filled, half-filled and empty in the ground state respectively.
Beginning at 2.2 eV are excitations which remove a single electron from the fully filled Cu 3d shell and add them to either the O 2p$_\pi$ or Cu 4s orbitals.
The neutral copper atom also carries a Hund's coupling between the d and s orbitals of about $\sim$ 0.5 eV, the effects of which would be measurable among the states in $\mathcal{LE}$ as long as states of both doublet and quartet states are included.
In order to probe all of these low-energy excitations we select a low-energy space
\begin{equation}
\mathcal{LE} = \text{Span(}\{ |\Psi \rangle | \langle \Psi | \hat{n}_{3d} | \Psi \rangle \ge 9,\ \hat{H}_\text{ab}|\Psi\rangle = E |\Psi\rangle \}\text{)}.
\label{eq:LE}
\end{equation}

Our sample space was generated by using multi-Slater-Jastrow trial wavefunctions in FN-DMC. 
In principle we could probe every low-energy excitation in $\mathcal{LE}$ in a statistically independent manner by just sampling the eigenstates within $\mathcal{LE}$ and including various non-eigenstates within $\mathcal{LE}$ as well to ensure robust statistics.
In practice we do not have access to these eigenstates and approximate methods are required to sample the low-energy space. 
Consider a trial wavefunction of the form $e^{J_i} |\text{D}_i\rangle$, where $e^{J_i}$ is a three-body Jastrow factor and $|\text{D}_i\rangle$ is a determinant of some single particle orbitals.
If the nodes of $|\text{D}_i\rangle$ are similar to the nodes of an eigenstate $|\Phi_i\rangle \in \mathcal{LE}$, then the properties and energy calculated from FN-DMC using this trial wavefunction will approximate those of the eigenstate, with the degree of approximation depending on the quality of the trial function nodal structure.
If we were able to find for each $|\Phi_i \rangle \in \mathcal{LE}$ such a $|\text{D}_i\rangle$ then we can approximate any state in $\mathcal{LE}$ as
\begin{equation}
|\Psi\rangle = \sum_i c_i |\Phi_i\rangle \sim \sum_i c_i \lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} (e^{J_i} |D_i\rangle) \xrightarrow{J_i = J} \lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} (e^{J}\sum_{i} c_i|\text{D}_i\rangle).
\label{eq:sampling}
\end{equation}
The last equality is valid if $\forall i\  J_i = J$, allowing us to map $|\Psi \rangle \in \mathcal{LE}$ to the end result of an FN-DMC calculation using a multi-Slater-Jastrow trial wavefunction. 
We use symmetry-targeted unrestricted Kohn-Sham (UKS) to generate our determinants $|\text{D}_i\rangle$ using a B3LYP functional, which has been shown to accurately reproduce ground state nodal properties of transition-metal oxide molecules, a Trail-Needs pseudopotential, and the VTZ Trail-Needs basis using the package PySCF.
This method allows us to access almost every excitation in $\mathcal{LE}$, however, there are some cases of two or more low-energy states within $\mathcal{LE}$ which have identical S$_z$ and number of electrons per irrep.
An example of this dilemma would be the ground state and the state with an excitation from Cu 3d$_{xz} \rightarrow $  O 2p$_x$.
In these few cases the higher energy excited $|\text{D}_i\rangle$ is inaccessible and the corresponding $|\Phi_i\rangle$ is excluded from the sum in \eqref{eq:sampling}, restricting our sample space to a subspace of $\mathcal{LE}$.
We will discuss the consequences of this restriction later on.
A single three-body Jastrow factor $J$ was used for every $|\text{D}_i\rangle$ and was optimized on the lowest energy UKS state using a linear energy optimization method.
The FN-DMC calculations were conducted with T-moves, to make the calculation variational while using pseudopotentials, with a timestep of $\tau = 0.01$.
The coefficients $\{c_j\}$ for our sample states were chosen via a shell-sampling method. 
We begin by fixing the coefficient $c_0 = \sqrt{w}$ where $w \in \{1.0, 0.8,..., 0.2\}$. 
For each choice of $w$ we sample n = 5 states by randomly selecting the unassigned coefficients from a uniform distribution such that $\sum_i c_i^2 = 1$. 
This procedure generates shells of samples at decreasing distance from the eigenstate $|\Phi_0\rangle$. 
We can then loop over all $i$ to generate shells near each eigenstate.

>>>>

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subfigure}
\usepackage[margin=1.0in]{geometry}
\usepackage{float}
\usepackage{ulem}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{wrapfig}
\renewcommand{\baselinestretch}{1.5}

\begin{document}
\section{Introduction}
A core problem of condensed matter physics is understanding how to systematically develop low-energy effective models for strongly correlated electron systems. In the forefront of modern research are materials like high temperature superconductors and heavy fermion systems for which simple effective theories have not yet been found. For example, the cuprate family of high-Tc superconductors have an intricate phase diagram in the doping-temperature space  with a robust anti-ferromagnetic (AFM) charge-transfer insulator state near zero doping, a prominent superconducting dome in the 0.1 - 0.2 hole doped region at low temperatures with psuedo-gap and non-Fermi liquid metallic phases at higher temperatures and Fermi liquid like behavior in the overdoped area, of which only the AFM insulating and Fermi liquid phases have well understood mechanism and models. The iron-pnictides also have superconducting domes, non-Fermi liquid like metallic phases, a spin-density-wave antiferommagnetic phase, all under various dopings and temperatures. Heavy-fermion materials have a purported quantum critical point between an AFM and Fermi-liquid phase which houses a superconducting dome, leading to another phase diagram for which building model Hamiltonians is very challenging.

The fundamental barrier in writing down low-energy theories for strong correlated materials is the existence of non-perturbative interactions between various degrees of freedom - for example, lattice, spin and charge. Consider just the cuprate case for now, the two phases which are well understood contain a singular important degree of freedom in the low-energy space. Near zero doping the cuprates are charge-transfer insulators with localized spin-1/2 low-energy degrees of freedom. In the heavily overdoped side, the lowest energy excitations look like well defined Fermi-liquid quasiparticles. In the intermediate doping region where the low-energy excitations cannot be easily sectioned into spin-like or electron-like only, we see the emergence of exotic phases and where most difficulties arise in writing down low-energy effective models. On top of that, this entire discussion did not include any lattice effects like structural phase transitions which many strongly-correlated materials have as well.

@@

We propose using a density matrix downfolding (DMD) procedure to systematically develop low-energy effective theories of the 1-band nearest-neighbor Hubbard model on a square lattice as a stepping stone towards building effective theories for strongly correlated materials. The Hubbard model is a simple model with nearest-neighbor hopping and a local on-site Coulomb repulsion, parameterized by a single ratio U/t: 
\begin{equation}
H_\text{hub} = -t \sum_{\langle i,j \rangle,\sigma}( c_{i^\dagger,\sigma} c_{j,\sigma} + h.c.) + U \sum_i n_{i\uparrow} n_{i\downarrow}
\label{hubbard}.
\end{equation}
The solutions in 1-d are well understood via the Bethe ansatz , however the 2-d square lattice Hubbard model has no known exact solution.  In the very strong coupling limit $U/t \rightarrow \infty$ an emergent low-energy spin-space exists with an effective nearest-neighbor AFM Heisenberg Hamiltonian. In the region $U = 0$ a trivial non-interacting theory holds. Between these two limits, in particular the intermediate coupling region $2 < U < 8$, no well established effective theory exists. In this region there is a metal-insulator transition at $U_c/t \sim 4$, and the pseudogap opens up at finite temperature around $U/t \sim 7$, leading to the development of a Mott gap for large $U/t > 11$. In addition to the half-filling phase diagram, this model exhibits a plethora of interesting ground state properties when doped including stripe phases and d-wave pairing. The relative simplicity of the Hubbard model and the lack of a well-established low-energy theory in a broad intermediate coupling region lends the model as a useful playground for our new downfolding procedure in preparation for tackling realistic strongly correlated electron systems.

The DMD procedure allows us to systematically develop low-energy effective theories for any Hamiltonian and Hilbert space $(H, \mathcal{H})$.
The procedure defines a method for downfolding the \textit{ab-initio} energy functional and Hilbert space onto an effective energy functional defined on a low-energy subspace of the Hilbert space 
\begin{equation}
(E, \mathcal{H}) \xrightarrow{\text{DMD}} (E_\text{eff}, \mathcal{LE} \subset \mathcal{H}), \text{ s.t. }
\forall |\Psi\rangle \in \mathcal{LE}, \ E_\text{eff}[\Psi] - E[\Psi] = \epsilon[\Psi] + E_0.
\label{eq:DMD}
\end{equation} 
The energy functionals can be written as expectations over Hamiltonian operators $E = \langle \Psi | \hat{H} |\Psi \rangle$, $E_\text{eff} = \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$, $\epsilon$ represents the error in our effective theory, and $E_0$ is a constant energy shift.
Broadly the DMD method consists of four steps which will be discussed in detail later: 
A) Defining the space of low-energy excitations which $E_\text{eff}$ should be able to describe accurately.
B) Sampling, at minimum, a set of states in $\mathcal{H}$ which probe the chosen low-energy excitations in a statistically independent manner. The selected states need not be eigenstates. 
C) Constructing a set of candidate models which are linear in reduced density matrix (RDM) descriptors with variable coefficients $\{c\}$
\begin{equation}
E_\text{eff} = \sum_k c_k \langle \Psi | \hat{d}_k |\Psi \rangle + E_0.
\label{eq:Eeff}
\end{equation}
Here $\hat{d}_k$ are Hermitian operators typically composed of 1- or 2-RDM elements. 
Examples of possible $\hat{d}_k$ are number operators $\hat{n}_i$ and exchange operators $\vec{S}_i \cdot \vec{S}_j$.
Additionally, a single particle basis must be built to express RDM elements on and should be chosen such that variations among states in $\mathcal{LE}$ are mostly described by variations in the projection of these states onto the basis.
D) Using the samples to fit the coefficients $\{c\}$ by linear regression such that $E_\text{eff}$ satisfies the conditions in \eqref{eq:DMD} with the smallest error $\epsilon$ possible. 
The fitting procedure is typically an ordinary linear regression, however, one may use a custom cost function in the regression as well. 
In order to conduct the linear regression for a functional like \eqref{eq:Eeff}, one needs to calculate for each state sampled from $\mathcal{LE}$ the expectation values of $\hat{H}$ and every descriptor $\{\hat{d}_k\}$ used in the regression.
In principle, if conducted with no approximations, the DMD method is rigorously guaranteed to obtain exact results, namely $\epsilon[\Psi] = 0$.
For real world systems we will work with $H = H_\text{ab}$, the \textit{ab-initio} Hamiltonian under the Born-Oppenheimer approximation with psuedopotentials:
\begin{equation}
\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|}
\label{eq:Hab}
\end{equation}
where $i$ indicates electrons and $I$ nuclei, and the unit of energy is Hartree (Ha).

When downfolding $\hat{H}_\text{ab}$ we will use FN-DMC to access states in $\mathcal{LE}$ and accurately calculate their \textit{ab-initio} energies and RDM elements.
Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function.
Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}_\text{ab}$ with ground state $|\Phi_0\rangle$. We apply the projector $e^{-\tau \hat{H}_\text{ab}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
\begin{equation}
\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} |\Psi_T\rangle 
\equiv \lim_{\tau \rightarrow \infty} |\Psi_\text{DMC}(\tau)\rangle \propto \langle \Phi_0|\Psi_T\rangle |\Phi_0\rangle,
\end{equation}
projecting out the ground state as long as the trial wave function we choose is not orthogonal to the ground state. 
The stochastic implementation involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H}_\text{ab} - E_T)} | R^\prime \rangle$. Since $H_\text{ab} = T + V$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(V(R) + V(R^\prime))/2} \langle R| e^{-d\tau(\hat{T} - E_T)}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.
This implementation suffers from a fermion sign problem which is dealt with through a fixed-node approximation where the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}_\text{ab}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.
We make use of this variational principle to access low-energy states within $\mathcal{H}$ by choosing trial wave functions with nodes that differ from the nodes of $|\Phi_0 \rangle$.
To calculate the expectation value of an operator $\hat{A}$ in FN-DMC we will use a mixed estimator $\langle \Psi_{DMC}(\tau) |\hat{A} | \Psi_T \rangle/\langle \Psi_\text{DMC}(\tau) | \Psi_T \rangle$ which allows us to use the importance sampled distribution $f = \Psi_\text{DMC}(\tau)\Psi_T$ for our Monte Carlo estimates.
Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.

@@

\subsection{Effective theory for CuO molecule using DMD}
As a first step towards developing accurate models for extended systems like solids using DMD, we constructed a many-body effective model for the CuO molecule which accurately describes the eigenstates and spectra seen in experiment up to 2eV above the ground state.
The neutral CuO molecule has been a subject of intense theoretical and experimental study due to the complex structure of its low-energy space.
While the low-energy excitations in the doublet sector of the CuO molecule are well understood, singlet selection rules for spectroscopic measurements and the doublet ground state leave the quartet sector of excitations unexplored.
We construct our $\mathcal{LE}$ based on recent anion photoelectron spectroscopy (APES) measurements of the low-lying excited states of the nuetral CuO molecule. 
The lowest lying excitations involve electrons in the Cu 3d, Cu 4s and O 2p orbitals.
Given that the lowest energy excitations contain a fully-filled Cu 3d shell or a single hole in this shell we define our low-energy space as:
\begin{equation}
\mathcal{LE} = \text{Span(}\{ |\Psi \rangle | \langle \Psi | \hat{n}_{3d} | \Psi \rangle \ge 9,\ \hat{H}_\text{ab}|\Psi\rangle = E |\Psi\rangle \}\text{)}.
\label{eq:LE}
\end{equation}

Our sample space was generated by using multi-Slater-Jastrow trial wavefunctions in FN-DMC.
The parameterization for our low-energy trial wave functions was
\begin{equation}
e^{J}\sum_{i} c_i|\text{D}_i\rangle = \sum_{i} c_i (e^J |\text{D}_i \rangle) \xrightarrow{FN-DMC} \ \sim \sum_i c_i |\Phi_i \rangle \in \mathcal{LE}
\label{eq:sampling}
\end{equation}
where $e^J$ is a three-body Jastrow factor and $|\text{D}_i\rangle$ is a Slater determinant whose nodes accurately approximate those of an eigenstate $|\Phi_i \rangle \in \mathcal{LE}$.
Under the FN-DMC projection these trial wave functions map to states which are very close to linear combinations of eigenstates in $\mathcal{LE}$.
We used symmetry-targeted unrestricted Kohn-Sham (UKS) to generate our determinants $|\text{D}_i\rangle$ using a B3LYP functional, a Trail-Needs pseudopotential, and the VTZ Trail-Needs basis using the package PySCF.
This method allows us to access almost every excitation in $\mathcal{LE}$, except the few cases when two states have identical S$_z$ and number of electrons per irrep.
An example would be the ground state and the state with an excitation from Cu 3d$_{xz} \rightarrow $  O 2p$_x$, where the latter would be inaccessible.
The FN-DMC calculations were conducted with T-moves to make the calculation variational while using pseudopotentials, and a timestep of $\tau = 0.01$.
The coefficients $\{c_j\}$ for our sample states were chosen via a shell-sampling method. 
We begin by fixing the coefficient $c_0 = \sqrt{w}$ where $w \in \{1.0, 0.8,..., 0.2\}$. 
For each choice of $w$ we sample n = 5 states by randomly selecting the unassigned coefficients from a uniform distribution such that $\sum_i c_i^2 = 1$. 
We can then loop over all $i$ to generate shells of decreasing distance near each eigenstate $|\Phi_i\rangle$.

Parameterizing our effective model requires selecting a set of descriptors from which we would like to build our model and a basis on which to build our descriptors.
Occupation energies of and hybridization between Cu 3d, Cu 4s and O 2p are necessary to describe the low-energy excitations in CuO and are best represented by the occupation energies in a molecular orbital basis (MO) which would neatly package both effects.
The lack of doubly occupied Cu 4s states in the low-energy spectrum of CuO and the existence of a large Hund's coupling between the Cu 4s and 3d orbitals in the Cu atom indicate the necessity for a Hubbard U$_s$ and Hund's J$_{sd}$ respectively. 
These two interaction terms are best represented in a localized intrinsic atomic orbital (IAO).
To account for orbital relaxation between excited states in the CuO molecule we will consider including additional hybridization between MOs.
In total our space of canditate models is then 
\begin{equation}
\begin{split}
\{\bar{n}_{4s}, \bar{n}_{p_\pi}, \bar{n}_{p_z}, \hat{n}_{4s\uparrow} \hat{n}_{4s\downarrow},\sum_{i \in \{xy, xz,...\}}\vec{S}_{4s}\cdot \vec{S}_{d_i}\} \ + \\
\text{P}(\bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{4s}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{4s} + h.c.)
\end{split}
\label{eq:models}
\end{equation}
where P denotes the power set and operators are defined on the IAO basis unless denoted by a bar, in which case the operator is built on the MO basis. 
The coefficients for each term above will be denoted $\bar{\epsilon}_{4s},\ \bar{\epsilon}_{p_\pi},\ \bar{\epsilon}_{p_z},\ U_s,\ J_{sd},\ \bar{t}_\pi,\ \bar{t}_{dz},\ \bar{t}_{sz},\ \bar{t}_{ds}$. 
All energies will be relative to $\bar{\epsilon_{3d}} = 0$. 
The symbol $\pi$ denotes a contraction over $x, y$ for p orbitals and $xz,\ yz$ for d orbitals. The symbol $\delta$, introduced later, denotes a contraction over $xy,\ x^2-y^2$ for d orbitals.

After fitting each of the potential models in \eqref{eq:models} using ordinary least squares (OLS) linear regression we find three models which describe the energy functional on our sample set accurately with low variance, but whose eigenstates and spectra differ as a consequence of intruder states below an energy of 2eV as illustrated in Figure \ref{fig:InitED}. 
All three models have a first excited state around 1eV, but only in the minimal model does this state fall near the sample set used for the fitting.
The large distance of the eigenstates for the other two models from our sample set indicates that their predicted eigenvalues may suffer from large extrapolation errors, and as such we will call them \textit{intruder} states.
In Figure \ref{fig:Intruder} we present all eigenstates among the three potential models which we classify as intruders using a k-nearest neighbor approach with k = 5.

\begin{figure}[H]
	\centering
	\subfigure[Results of exact diagonalization showing the energies and properties of the first twenty eigenstates for three candidate models fit using ordinary linear regression. The ground state and first excited state are colored and have enlarged markers for clarity. Error bars are 95\% CI calculated by a bootstrap estimate.]{
	 	\label{fig:InitED}
		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/init_ed.pdf}
	}
	\qquad
	\subfigure[Collection of unique intruder states from our three potential models selected using a k-nearest neighbor approach with k = 5.]{	
		\label{fig:Intruder}
		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/intruder.pdf}
	}
	\caption{Results of OLS fitting on sampled data and selected intruder states.}
\end{figure}

Given that our intruder states correspond to states within $\mathcal{LE}$ but outside our sample set, they must live within the span of the eigenstates which were excluded in our sampling scheme as described before.
This is evidenced by the fact that two of the intruder states correspond explicitly to excitations we could not access: Cu 3d$_{z^2} \rightarrow $ O 2p$_\pi$ and  Cu 3d$_\pi \rightarrow $ O 2p$_\pi$.
We can approximate these states by single rigid MO excitations above the UKS ground state, which results in states with a FN-DMC energy above 4eV.
Including a generous 2eV reduction in FN-DMC energy due to orbital relaxation, we estimate that these two states should lie at least 2eV in energy. 
Since all the other inaccessible eigenstates are double or higher order excitations, we assert a prior: intruder states should lie above 2eV in energy in our effective theories, enforced via an augmented cost function
\begin{equation}
\text{Cost} = \sum_{i} (E_\text{eff}[\Psi_i] - E_\text{ab}[\Psi_i])^2 + \lambda \sum_{p}\text{QHL}(2 - E_\text{eff}[\Psi_p]),\ \text{QHL}(x) = \Theta(x)x^2
\label{eq:cost}
\end{equation}
where $\lambda>0$ is a parameter which can be varied, QHL is a quadratic hinge loss, $\Theta$ is the Heaviside step function, the index $i$ is over our sampled states and $p$ over the selected intruder states.

\begin{figure}[H]
	\centering
	\subfigure[On the left, scores of the three candidate models at various $\lambda$ when fitting our effective theories using the cost function \eqref{eq:cost}. On the right we show predicted versus calculated energies for our low-energy sampled states for the minimal model fit at $\lambda = 20$.]{
	 	\label{fig:Prior}
		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/prior_and_regr.pdf}
	}
	\qquad
	\subfigure[Results of exact diagonalization showing the energies and properties of the first thirty eigenstates for three candidate models using linear regression with cost function \ref{eq:cost} and $\lambda = 20$. The ground state and first excited state are colored and have enlarged markers for clarity. Error bars are 95\% CI calculated by a bootstrap estimate.]{	
		\label{fig:FinalED}
		\includegraphics[width=0.45\textwidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/final_ed.pdf}
	}
	\caption{Results of linear regression using augmented cost function \eqref{eq:cost}.}
\end{figure}

Using our new cost function we find a set of models which accurately describe our sample data, have very similar spectra and eigenvectors to eachother and experiment, and do not contain intruder states below the 2eV prior. 
Shown in Figure \ref{fig:Prior} are the R$^2$ scores and QHL for the three candidate models fit at $\lambda \in [0,20]$. 
We believe that all three models at $\lambda = 20$ should be considered good models as they satisfy our prior and also explain our sample set accurately. 
Also shown are $E_\text{eff}, E_\text{ab}$ for our samples using the minimal model at $\lambda = 20$ to illustrate the high fit quality even with a large $\lambda$.
Figure \ref{fig:FinalED} presents the properties and eigenvalues of the lowest 30 eigenstates for the three fit models at $\lambda = 20$. 
The additional constraint from asserting a prior has seemingly caused a single effective theory to emerge from our set of candidate models.
This conclusion is verified by looking at the values of the fit parameters for our three models, presented in Table 1.
All occupation energies are relative to $\epsilon_{d_\delta}$.
In the doublet sector the ground and first excited state of any of the three models match experiment, as well as the block of states at 2eV corresponding to excitations out of the Cu 3d shell.
The lowest energy quartet states are at 1, 2eV respectively and are in line with the observation of a quartet state at 2eV in APES measurements.
